### 2017-04-30 sec-limitations.tex
Revised limitations.

There are a few limitations of our model that are important to highlight.
First, there are specific technical limitations due to the statistical
nature of the model and the particular methodology we propose for fitting
our model parameters. Second, there are limitations in the kinds of
patterns our model is able to discover in its current formulation, and the
ease with which instructors are able to extract knowledge from these
patterns. We address both of these lines below.

\subsection{Technical Limitations and Implementation Challenges}
In particular, because the model is complex and has many parameters,
to truly uncover relevant patterns in the data, a large amount of it must
be available to the training algorithm. The assumption that we have a large
amount of sequence data available for training on should generally hold for
most MOOC courses, but this assumption may be problematic if attempting to
apply our model to smaller online (or on-campus) courses.

Our model fits its parameters using maximum likelihood estimation using the
EM algorithm. The EM algorithm is a hill climbing algorithm that is
optimizing in a highly non-convex parameter space. Thus, it can only
guarantee reaching a local maximum in practice~\citep{Dempster:1977:JRSS}.
This may mean that the parameters found for the model may not be the
``best'' parameters in a global sense, which may lead to suboptimal latent
state representations and transition patterns. A commonly applied approach
to address this drawback is to run the model multiple times with different
starting points to allow the model to explore a larger portion of the
potential parameter space. One can then compare the log-likelihood of the
data between the models that were started from different initialization
points and select the one that has the highest value. This still does not
guarantee finding a global maximum, but it does help alleviate the
potential for finding a particularly bad local maximum. In practice, we
have found that our model can be fit to the data fairly quickly on
commodity hardware, so running it multiple times to address this concern is
not computationally unfeasible.

@@

@@

While the model can discover \emph{patterns} in the data in a fully
automated way, there is still clearly a burden on the user of our model to
interpret the patterns it has discovered to create actionable
\emph{knowledge} about the MOOC from which the data was extracted.
Extracting the patterns is a necessary step towards the creation of
knowledge, and we view our model as a component in a collaborative system
which leverages the machine to perform statistical modeling to extract
patterns which then enable a human actor to extract knowledge and take
actions on the basis of the data. The pattern discovery is an important
and absolutely necessary component, but it is not the end of the knowledge
discovery process.

>>>>

There are a few limitations of our model that are important to highlight.
First, there are specific technical limitations due to the statistical
nature of the model and the particular methodology we propose for fitting
our model parameters. Second, there are limitations in the kinds of
patterns our model is able to discover in its current formulation, and the
ease with which instructors are able to extract knowledge from these
patterns. We discuss both of these lines below.

\subsection{Technical Limitations and Implementation Challenges}

 %irregular behavior
%patterns In particular, because
One potential limitation is that the model is complex and has many parameters,
to truly uncover relevant patterns in the data, thus a large amount of it must
be available to the training algorithm. The assumption that we have a large
amount of sequence data available for training on should generally hold for
most MOOC courses, but this assumption may be problematic if attempting to
apply our model to smaller online (or on-campus) courses.

Our model fits its parameters using maximum likelihood estimation using the
EM algorithm. The EM algorithm is a hill climbing algorithm that is
optimizing in a highly non-convex parameter space. Thus, it can only
guarantee reaching a local maximum in practice~\citep{Dempster:1977:JRSS}.
This may mean that the parameters found for the model may not be the
``best'' parameters in a global sense, which may lead to suboptimal latent
state representations and transition patterns. The problem may not be as serious as it 
appears to be because strong patterns tend to always show up as long as the algorithm reaches a reasonably good local maximum, and the differences of the results tend to be 
related to weak ``unstable'' patterns which may not be reliable anyway. 
 A commonly applied approach
to address the problem of multiple local maxima is to run the model multiple times with different
starting points to allow the model to explore a larger portion of the
potential parameter space. One can then compare the log-likelihood of the
data between the models that were started from different initialization
points and select the one that has the highest value. This still does not
guarantee finding a global maximum, but it does help alleviate the
potential for finding a particularly bad local maximum.  
In practice, we
have found that our model can be fit to the data fairly quickly on
commodity hardware, so running it multiple times to address this concern is
not computationally unfeasible.

@@

The proposed behavior representation is most suitable for representing
recurring behavior patterns, which presumably are most interesting 
to extract from the data, but may not cover all the interesting patterns in the data. 
It would be interesting to further explore other complementary models such as
 time series models, which may help capture non-recurring patterns.

@@

While the model can discover \emph{patterns} in the data in a fully
automated way, there is still clearly a burden on the user of our model to
interpret the patterns it has discovered to create actionable
\emph{knowledge} about the MOOC from which the data was extracted.
Extracting the patterns is a necessary step towards the creation of
knowledge, and we view our model as a component in a collaborative system
which leverages the machine to perform statistical modeling to extract
patterns which then enable a human actor to extract knowledge and take
actions on the basis of the data. The pattern discovery is an important
and absolutely necessary component, without which it would be very difficult, if not impossible, 
for humans to directly digest the student behavior buried in the large amount of 
data.  Of course, pattern discovery is only a means to help humans obtain knowledge, not the end of the knowledge discovery process.
